CF Registry Analysis

### 2024-01-31 Christine Chan/CF Registry/cf_registry_analysis.qmd
cf registry time varying covariates fixed

# CFRD incidence and prevalence

```{r}
# All data
encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
  na.strings = c("", "N/A")
)
annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual <- annual %>% arrange(eDWID, ReviewYear)
# Add age
ages <- annual %>%
  mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>%
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID, reviewyear) %>%
  summarise(
    `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`),
      decreasing = TRUE
    )[1])),
    Age = Age_YrEnd[1], .groups = "drop"
  )
encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear"))
encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`,
  c(-Inf, 10, 19, 31, Inf),
  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
  right = F
)
encounter$`Age Group` <- cut(encounter$Age,
  c(-Inf, 10, 19, 31, Inf),
  labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
  right = F
)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>%
  filter(cfrd_status == 3) %>%
  group_by(eDWID) %>%
  filter(row_number() == 1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_cffpr,
  by = c("eDWID", "Age in 2018")
)
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
# Ordered factors
encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr,
  levels = c("No", "Yes"), ordered = T
)
# Remove those missing age
encounter <- encounter %>% filter(!is.na(Age))
```

```{r}
#| label: tbl-incidence_cffpr
#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
# For each year, pull everyone with and without CFRD that year
year_inc <- function(df = encounter, year, py = 1000) {
  # Filter and count
  inc <- df %>%
    filter(reviewyear == year, is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>%
    group_by(eDWID) %>%
    filter(row_number() == 1) %>%
    select(eDWID, `Age Group`, cfrd_cffpr)
  inc <- inc %>%
    group_by(cfrd_cffpr) %>%
    count(`Age Group`, .drop = F)
  inc <- inc[!is.na(inc$`Age Group`), ]
  # Calculations
  new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"])
  total_at_risk <- sum(inc$n)
  total_inc <- new_cases / (total_at_risk / py)

  new_cases_10 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"])
  at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"])
  inc_10 <- new_cases_10 / (at_risk_10 / py)

  new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "10 - 18"])
  at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)

  new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "19 - 30"])
  at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)

  new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "> 30"])
  at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
  inc_30 <- new_cases_30 / (at_risk_30 / py)

  # Nice results
  list(
    `Year` = year,
    `Total New Cases` = new_cases,
    `Total At Risk` = total_at_risk,
    `Total Incidence` = total_inc,
    `Total New Cases Age < 10` = new_cases_10,
    `Total At Risk Age < 10` = at_risk_10,
    `Total Incidence Age < 10` = inc_10,
    `Total New Cases Age 10-18` = new_cases_10_18,
    `Total At Risk Age 10-18` = at_risk_10_18,
    `Total Incidence Age 10-18` = inc_10_18,
    `Total New Cases Age 19-30` = new_cases_19_30,
    `Total At Risk Age 19-30` = at_risk_19_30,
    `Total Incidence Age 19-30` = inc_19_30,
    `Total New Cases Age > 30` = new_cases_30,
    `Total At Risk Age > 30` = at_risk_30,
    `Total Incidence Age > 30` = inc_30
  )
}
incidence <- lapply(2003:2018, function(y) {
  year_inc(year = y)
})
incidence <- data.frame(do.call(rbind, incidence), check.names = F)
incidence <- data.frame(lapply(incidence, as.numeric), check.names = F)
kable(incidence, digits = 3)
```

In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases.

```{r}
#| label: tbl-prevalence_cffpr
#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
encounter %>%
  group_by(eDWID, reviewyear) %>%
  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
  group_by(reviewyear) %>%
  summarise(
    Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100,
    .groups = "drop"
  ) %>%
  kable(., digits = 3)
```

```{r}
#| label: tbl-prevalence_cffpr_year
#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone"
encounter %>%
  group_by(eDWID, reviewyear, `Age Group`) %>%
  summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>%
  group_by(reviewyear, `Age Group`) %>%
  summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>%
  pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>%
  kable(., digits = 3)
```

In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.

# CFRD screening rates

```{r}
#| label: tbl-screening_ogtt
#| tbl-cap: "Yearly OGTT Screening Rates"
year_screen <- function(df = encounter, year) {
  # Filter and count
  inc <- df %>%
    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
    mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>%
    group_by(eDWID) %>%
    summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1])
  inc <- inc %>%
    group_by(ogtt) %>%
    count(`Age Group`) %>%
    filter(`Age Group` != "< 10")
  # Calculations
  total_screened <- sum(inc$n[inc$ogtt == 1])
  total_seen <- sum(inc$n)
  screen_rate <- total_screened / total_seen * 100

  screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"]
  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
  rate_10_18 <- screened_10_18 / seen_10_18 * 100

  screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"]
  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
  rate_19_30 <- screened_19_30 / seen_19_30 * 100

  screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"]
  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
  rate_30 <- screened_30 / seen_30 * 100

  # Nice results
  list(
    `Year` = year,
    `Total Screened` = total_screened,
    `Total Seen` = total_seen,
    `Screen Rate (%)` = screen_rate,
    `Total Screened Age 10-18` = screened_10_18,
    `Total Seen Age 10-18` = seen_10_18,
    `Screen Rate Age 10-18 (%)` = rate_10_18,
    `Total Screened Age 19-30` = screened_19_30,
    `Total Seen Age 19-30` = seen_19_30,
    `Screen Rate Age 19-30 (%)` = rate_19_30,
    `Total Screened Age > 30` = screened_30,
    `Total Seen Age > 30` = seen_30,
    `Screen Rate Age > 30 (%)` = rate_30
  )
}
screen <- lapply(2003:2018, function(y) {
  year_screen(year = y)
})
screen <- data.frame(do.call(rbind, screen), check.names = F)
screen <- data.frame(lapply(screen, as.numeric), check.names = F)
kable(screen, digits = 3)
```

```{r}
#| label: tbl-screening_a1c
#| tbl-cap: "Yearly HbA1c Screening Rates"
year_screen <- function(df = encounter, year) {
  # Filter and count
  inc <- df %>%
    filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>%
    mutate(a1c = !is.na(hgba1c)) %>%
    group_by(eDWID) %>%
    summarise(a1c = max(a1c), `Age Group` = `Age Group`[1])
  inc <- inc %>%
    group_by(a1c) %>%
    count(`Age Group`) %>%
    filter(`Age Group` != "< 10")
  # Calculations
  total_screened <- sum(inc$n[inc$a1c == 1])
  total_seen <- sum(inc$n)
  screen_rate <- total_screened / total_seen * 100

  screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"]
  seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"])
  rate_10_18 <- screened_10_18 / seen_10_18 * 100

  screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"]
  seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"])
  rate_19_30 <- screened_19_30 / seen_19_30 * 100

  screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"]
  seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"])
  rate_30 <- screened_30 / seen_30 * 100

  # Nice results
  list(
    `Year` = year,
    `Total Screened` = total_screened,
    `Total Seen` = total_seen,
    `Screen Rate (%)` = screen_rate,
    `Total Screened Age 10-18` = screened_10_18,
    `Total Seen Age 10-18` = seen_10_18,
    `Screen Rate Age 10-18 (%)` = rate_10_18,
    `Total Screened Age 19-30` = screened_19_30,
    `Total Seen Age 19-30` = seen_19_30,
    `Screen Rate Age 19-30 (%)` = rate_19_30,
    `Total Screened Age > 30` = screened_30,
    `Total Seen Age > 30` = seen_30,
    `Screen Rate Age > 30 (%)` = rate_30
  )
}
screen <- lapply(2003:2018, function(y) {
  year_screen(year = y)
})
screen <- data.frame(do.call(rbind, screen), check.names = F)
screen <- data.frame(lapply(screen, as.numeric), check.names = F)
kable(screen, digits = 3)
```

In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded. Screening rate is calculated as the number of people with an OGTT or HbA1c, respectively, divided by the total number of people seen in clinic that year.

@@

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc", "MutClass", "A_IsOnEnzymes",
  "A_FEV1_pct_predicted", "tube_feeding", "NumPulmExacerbation",
  "A_pseudomonasaeruginosa", "A_fungalyeast1", "A_mycocultureresults3",
  "A_pulmonarycomplications1", "A_corticosteroids1", "cfrd_status_annual",
  "A_Vx770", "A_VX809comb", "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  A_FEV1_pct_predicted = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_pulmonarycomplications1 = "ABPA",
  A_mycocultureresults3 = "Microorganisms in myco culture"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd,
    Age_YrEnd >= 10
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(
    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
    age_at_entry, everything()
  )
# Convert from tibble to dataframe for survival packages
s <- as.data.frame(s)
# CFRD status to numeric
s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
# Drop under 10s
s$age_group <- droplevels(s$age_group)
```

@@

```{r}
# Whole cohort
km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
  data = s, cluster = eDWID
)
ggsurvplot(km_fit,
  data = s, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018), break.x.by = 1
)
kable(tidy(km_fit))
# By age group
km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
  data = s, cluster = eDWID
)
ggsurvplot(km_fit,
  data = s, risk.table = T, cumevents = T,
  fun = "event", xlim = c(2003, 2018), break.x.by = 1
)
km_table <- tidy(km_fit)
km_table <- split.data.frame(km_table, km_table$strata)
invisible(lapply(km_table, function(t) {
  name <- t$strata[1]
  t <- t %>% select(-strata)
  kable(t, caption = name)
}))
```

@@

```{r}
s_split <- survSplit(Surv(tstart, tstop, cfrd_status_annual) ~ .,
  data = s, cut = 2003:2018, episode = "tgroup"
)
cox_res_split <- coxph(update(f, . ~ Gender:strata(tgroup)), data = s_split,
                       cluster = eDWID)
t = tidy(cox_res_split,exponentiate = T)
```

#### P. aeruginosa

>>>>

<!-- # CFRD incidence and prevalence -->

<!-- ```{r} -->
<!-- # All data -->
<!-- encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv", -->
<!--   na.strings = c("", "N/A") -->
<!-- ) -->
<!-- annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv", -->
<!--   na.strings = c("", "N/A") -->
<!-- ) -->
<!-- encounter <- encounter %>% -->
<!--   arrange(eDWID, reviewyear, encounternum) -->
<!-- annual <- annual %>% arrange(eDWID, ReviewYear) -->
<!-- # Add age -->
<!-- ages <- annual %>% -->
<!--   mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd)) %>% -->
<!--   rename(reviewyear = ReviewYear) %>% -->
<!--   group_by(eDWID, reviewyear) %>% -->
<!--   summarise( -->
<!--     `Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), -->
<!--       decreasing = TRUE -->
<!--     )[1])), -->
<!--     Age = Age_YrEnd[1], .groups = "drop" -->
<!--   ) -->
<!-- encounter <- left_join(encounter, ages, by = c("eDWID", "reviewyear")) -->
<!-- encounter$`Age Group 2018` <- cut(encounter$`Age in 2018`, -->
<!--   c(-Inf, 10, 19, 31, Inf), -->
<!--   labels = c("< 10", "10 - 18", "19 - 30", "> 30"), -->
<!--   right = F -->
<!-- ) -->
<!-- encounter$`Age Group` <- cut(encounter$Age, -->
<!--   c(-Inf, 10, 19, 31, Inf), -->
<!--   labels = c("< 10", "10 - 18", "19 - 30", "> 30"), -->
<!--   right = F -->
<!-- ) -->
<!-- # For those with CFRD by registry criteria, find first visit with diagnosis -->
<!-- first_cfrd_cffpr <- encounter %>% -->
<!--   filter(cfrd_status == 3) %>% -->
<!--   group_by(eDWID) %>% -->
<!--   filter(row_number() == 1) %>% -->
<!--   select(eDWID, reviewyear, `Age in 2018`) %>% -->
<!--   rename(first_cfrd_cffpr = reviewyear) -->
<!-- # Add to encounter level data -->
<!-- encounter <- left_join(encounter, first_cfrd_cffpr, -->
<!--   by = c("eDWID", "Age in 2018") -->
<!-- ) -->
<!-- # Fill down CFRD Yes/No by year -->
<!-- encounter <- encounter %>% -->
<!--   group_by(eDWID) %>% -->
<!--   mutate( -->
<!--     cfrd_cffpr = case_when( -->
<!--       reviewyear < first_cfrd_cffpr ~ "No", -->
<!--       reviewyear >= first_cfrd_cffpr ~ "Yes" -->
<!--     ) -->
<!--   ) %>% -->
<!--   ungroup() -->
<!-- encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No" -->
<!-- # Ordered factors -->
<!-- encounter$cfrd_cffpr <- factor(encounter$cfrd_cffpr, -->
<!--   levels = c("No", "Yes"), ordered = T -->
<!-- ) -->
<!-- # Remove those missing age -->
<!-- encounter <- encounter %>% filter(!is.na(Age)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: tbl-incidence_cffpr -->
<!-- #| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone" -->
<!-- # For each year, pull everyone with and without CFRD that year -->
<!-- year_inc <- function(df = encounter, year, py = 1000) { -->
<!--   # Filter and count -->
<!--   inc <- df %>% -->
<!--     filter(reviewyear == year, is.na(first_cfrd_cffpr) | first_cfrd_cffpr >= year) %>% -->
<!--     group_by(eDWID) %>% -->
<!--     filter(row_number() == 1) %>% -->
<!--     select(eDWID, `Age Group`, cfrd_cffpr) -->
<!--   inc <- inc %>% -->
<!--     group_by(cfrd_cffpr) %>% -->
<!--     count(`Age Group`, .drop = F) -->
<!--   inc <- inc[!is.na(inc$`Age Group`), ] -->
<!--   # Calculations -->
<!--   new_cases <- sum(inc$n[inc$cfrd_cffpr == "Yes"]) -->
<!--   total_at_risk <- sum(inc$n) -->
<!--   total_inc <- new_cases / (total_at_risk / py) -->

<!--   new_cases_10 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "< 10"]) -->
<!--   at_risk_10 <- sum(inc$n[inc$`Age Group` == "< 10"]) -->
<!--   inc_10 <- new_cases_10 / (at_risk_10 / py) -->

<!--   new_cases_10_18 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "10 - 18"]) -->
<!--   at_risk_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
<!--   inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py) -->

<!--   new_cases_19_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "19 - 30"]) -->
<!--   at_risk_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
<!--   inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py) -->

<!--   new_cases_30 <- sum(inc$n[inc$cfrd_cffpr == "Yes" & inc$`Age Group` == "> 30"]) -->
<!--   at_risk_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
<!--   inc_30 <- new_cases_30 / (at_risk_30 / py) -->

<!--   # Nice results -->
<!--   list( -->
<!--     `Year` = year, -->
<!--     `Total New Cases` = new_cases, -->
<!--     `Total At Risk` = total_at_risk, -->
<!--     `Total Incidence` = total_inc, -->
<!--     `Total New Cases Age < 10` = new_cases_10, -->
<!--     `Total At Risk Age < 10` = at_risk_10, -->
<!--     `Total Incidence Age < 10` = inc_10, -->
<!--     `Total New Cases Age 10-18` = new_cases_10_18, -->
<!--     `Total At Risk Age 10-18` = at_risk_10_18, -->
<!--     `Total Incidence Age 10-18` = inc_10_18, -->
<!--     `Total New Cases Age 19-30` = new_cases_19_30, -->
<!--     `Total At Risk Age 19-30` = at_risk_19_30, -->
<!--     `Total Incidence Age 19-30` = inc_19_30, -->
<!--     `Total New Cases Age > 30` = new_cases_30, -->
<!--     `Total At Risk Age > 30` = at_risk_30, -->
<!--     `Total Incidence Age > 30` = inc_30 -->
<!--   ) -->
<!-- } -->
<!-- incidence <- lapply(2003:2018, function(y) { -->
<!--   year_inc(year = y) -->
<!-- }) -->
<!-- incidence <- data.frame(do.call(rbind, incidence), check.names = F) -->
<!-- incidence <- data.frame(lapply(incidence, as.numeric), check.names = F) -->
<!-- kable(incidence, digits = 3) -->
<!-- ``` -->

<!-- In @tbl-incidence_cffpr, incidence per year is calculated as the number of new CFRD diagnoses (by registry definition alone) per 1,000 person-years of those at risk, excluding previously identified cases. -->

<!-- ```{r} -->
<!-- #| label: tbl-prevalence_cffpr -->
<!-- #| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone" -->
<!-- encounter %>% -->
<!--   group_by(eDWID, reviewyear) %>% -->
<!--   summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>% -->
<!--   group_by(reviewyear) %>% -->
<!--   summarise( -->
<!--     Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, -->
<!--     .groups = "drop" -->
<!--   ) %>% -->
<!--   kable(., digits = 3) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: tbl-prevalence_cffpr_year -->
<!-- #| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone" -->
<!-- encounter %>% -->
<!--   group_by(eDWID, reviewyear, `Age Group`) %>% -->
<!--   summarise(cfrd = max(cfrd_cffpr), .groups = "drop") %>% -->
<!--   group_by(reviewyear, `Age Group`) %>% -->
<!--   summarise(Prevalence = mean(as.numeric(cfrd) - 1, na.rm = T) * 100, .groups = "drop") %>% -->
<!--   pivot_wider(names_from = `Age Group`, values_from = Prevalence) %>% -->
<!--   kable(., digits = 3) -->
<!-- ``` -->

<!-- In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year. -->

<!-- # CFRD screening rates -->

<!-- ```{r} -->
<!-- #| label: tbl-screening_ogtt -->
<!-- #| tbl-cap: "Yearly OGTT Screening Rates" -->
<!-- year_screen <- function(df = encounter, year) { -->
<!--   # Filter and count -->
<!--   inc <- df %>% -->
<!--     filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>% -->
<!--     mutate(ogtt = !is.na(ogttfast_bloodglucose) | !is.na(twohour_bloodglucose)) %>% -->
<!--     group_by(eDWID) %>% -->
<!--     summarise(ogtt = max(ogtt), `Age Group` = `Age Group`[1]) -->
<!--   inc <- inc %>% -->
<!--     group_by(ogtt) %>% -->
<!--     count(`Age Group`) %>% -->
<!--     filter(`Age Group` != "< 10") -->
<!--   # Calculations -->
<!--   total_screened <- sum(inc$n[inc$ogtt == 1]) -->
<!--   total_seen <- sum(inc$n) -->
<!--   screen_rate <- total_screened / total_seen * 100 -->

<!--   screened_10_18 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "10 - 18"] -->
<!--   seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
<!--   rate_10_18 <- screened_10_18 / seen_10_18 * 100 -->

<!--   screened_19_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "19 - 30"] -->
<!--   seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
<!--   rate_19_30 <- screened_19_30 / seen_19_30 * 100 -->

<!--   screened_30 <- inc$n[inc$ogtt == 1 & inc$`Age Group` == "> 30"] -->
<!--   seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
<!--   rate_30 <- screened_30 / seen_30 * 100 -->

<!--   # Nice results -->
<!--   list( -->
<!--     `Year` = year, -->
<!--     `Total Screened` = total_screened, -->
<!--     `Total Seen` = total_seen, -->
<!--     `Screen Rate (%)` = screen_rate, -->
<!--     `Total Screened Age 10-18` = screened_10_18, -->
<!--     `Total Seen Age 10-18` = seen_10_18, -->
<!--     `Screen Rate Age 10-18 (%)` = rate_10_18, -->
<!--     `Total Screened Age 19-30` = screened_19_30, -->
<!--     `Total Seen Age 19-30` = seen_19_30, -->
<!--     `Screen Rate Age 19-30 (%)` = rate_19_30, -->
<!--     `Total Screened Age > 30` = screened_30, -->
<!--     `Total Seen Age > 30` = seen_30, -->
<!--     `Screen Rate Age > 30 (%)` = rate_30 -->
<!--   ) -->
<!-- } -->
<!-- screen <- lapply(2003:2018, function(y) { -->
<!--   year_screen(year = y) -->
<!-- }) -->
<!-- screen <- data.frame(do.call(rbind, screen), check.names = F) -->
<!-- screen <- data.frame(lapply(screen, as.numeric), check.names = F) -->
<!-- kable(screen, digits = 3) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: tbl-screening_a1c -->
<!-- #| tbl-cap: "Yearly HbA1c Screening Rates" -->
<!-- year_screen <- function(df = encounter, year) { -->
<!--   # Filter and count -->
<!--   inc <- df %>% -->
<!--     filter(reviewyear == year, first_cfrd_cffpr > year | is.na(first_cfrd_cffpr)) %>% -->
<!--     mutate(a1c = !is.na(hgba1c)) %>% -->
<!--     group_by(eDWID) %>% -->
<!--     summarise(a1c = max(a1c), `Age Group` = `Age Group`[1]) -->
<!--   inc <- inc %>% -->
<!--     group_by(a1c) %>% -->
<!--     count(`Age Group`) %>% -->
<!--     filter(`Age Group` != "< 10") -->
<!--   # Calculations -->
<!--   total_screened <- sum(inc$n[inc$a1c == 1]) -->
<!--   total_seen <- sum(inc$n) -->
<!--   screen_rate <- total_screened / total_seen * 100 -->

<!--   screened_10_18 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "10 - 18"] -->
<!--   seen_10_18 <- sum(inc$n[inc$`Age Group` == "10 - 18"]) -->
<!--   rate_10_18 <- screened_10_18 / seen_10_18 * 100 -->

<!--   screened_19_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "19 - 30"] -->
<!--   seen_19_30 <- sum(inc$n[inc$`Age Group` == "19 - 30"]) -->
<!--   rate_19_30 <- screened_19_30 / seen_19_30 * 100 -->

<!--   screened_30 <- inc$n[inc$a1c == 1 & inc$`Age Group` == "> 30"] -->
<!--   seen_30 <- sum(inc$n[inc$`Age Group` == "> 30"]) -->
<!--   rate_30 <- screened_30 / seen_30 * 100 -->

<!--   # Nice results -->
<!--   list( -->
<!--     `Year` = year, -->
<!--     `Total Screened` = total_screened, -->
<!--     `Total Seen` = total_seen, -->
<!--     `Screen Rate (%)` = screen_rate, -->
<!--     `Total Screened Age 10-18` = screened_10_18, -->
<!--     `Total Seen Age 10-18` = seen_10_18, -->
<!--     `Screen Rate Age 10-18 (%)` = rate_10_18, -->
<!--     `Total Screened Age 19-30` = screened_19_30, -->
<!--     `Total Seen Age 19-30` = seen_19_30, -->
<!--     `Screen Rate Age 19-30 (%)` = rate_19_30, -->
<!--     `Total Screened Age > 30` = screened_30, -->
<!--     `Total Seen Age > 30` = seen_30, -->
<!--     `Screen Rate Age > 30 (%)` = rate_30 -->
<!--   ) -->
<!-- } -->
<!-- screen <- lapply(2003:2018, function(y) { -->
<!--   year_screen(year = y) -->
<!-- }) -->
<!-- screen <- data.frame(do.call(rbind, screen), check.names = F) -->
<!-- screen <- data.frame(lapply(screen, as.numeric), check.names = F) -->
<!-- kable(screen, digits = 3) -->
<!-- ``` -->

<!-- In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by registry definition) and those under age 10 were excluded. Screening rate is calculated as the number of people with an OGTT or HbA1c, respectively, divided by the total number of people seen in clinic that year. -->

@@

```{r}
# List of covariates
vars <- c(
  "age_group", "Gender", "bmi_perc_cat", "MutClass", "A_IsOnEnzymes",
  "ppFEV1_cat", "tube_feeding", "NumPulmExacerbation",
  "A_pseudomonasaeruginosa", "A_fungalyeast1", "A_mycocultureresults3",
  "A_pulmonarycomplications1", "A_corticosteroids1", "cfrd_status_annual",
  "A_Vx770", "A_VX809comb", "liver_disease"
)
# Nice labels for plots and tables
labels <- list(
  cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
  Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
  ppFEV1_cat = "FEV1 percent predicted",
  tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
  A_corticosteroids1 = "Use of corticosteroids",
  NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
  height_perc = "Height percentile", bmi_perc_cat = "BMI percentile",
  A_Vx770 = "Use of CFTR modulator Vx770",
  A_VX809comb = "Use of CFTR modulator Vx809 combination",
  A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
  A_fungalyeast1 = "Aspergillus (any species)",
  A_pulmonarycomplications1 = "ABPA",
  A_mycocultureresults3 = "Microorganisms in myco culture"
)
# For now, let's just assume that everyone's first CFRD+ diagnosis is correct
s <- annual %>%
  filter(
    !is.na(cfrd_status_annual),
    Age_YrEnd != 0,
    is.na(first_cfrd) | ReviewYear <= first_cfrd,
    Age_YrEnd >= 10
  ) %>%
  select(
    eDWID, ReviewYear, birth_year, Age_YrEnd,
    all_of(vars), cfrd_status_annual
  ) %>%
  rename(tstop = ReviewYear) %>%
  group_by(eDWID) %>%
  mutate(
    tstart = lag(tstop, default = birth_year[1]),
    age_at_entry = Age_YrEnd[1]
  ) %>%
  select(
    eDWID, birth_year, tstart, tstop, cfrd_status_annual,
    age_at_entry, everything()
  )
# Convert from tibble to dataframe for survival packages
s <- as.data.frame(s)
# CFRD status to numeric
s$cfrd_status_annual <- as.numeric(s$cfrd_status_annual) - 1
# Drop under 10s
s$age_group <- droplevels(s$age_group)
```

@@

```{r}
# # Whole cohort
# km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ 1,
#   data = s, cluster = eDWID
# )
# ggsurvplot(km_fit,
#   data = s, risk.table = T, cumevents = T,
#   fun = "event", xlim = c(2003, 2018), break.x.by = 1
# )
# kable(tidy(km_fit))
# # By age group
# km_fit <- survfit(Surv(tstart, tstop, cfrd_status_annual) ~ age_group,
#   data = s, cluster = eDWID
# )
# ggsurvplot(km_fit,
#   data = s, risk.table = T, cumevents = T,
#   fun = "event", xlim = c(2003, 2018), break.x.by = 1
# )
# km_table <- tidy(km_fit)
# km_table <- split.data.frame(km_table, km_table$strata)
# invisible(lapply(km_table, function(t) {
#   name <- t$strata[1]
#   t <- t %>% select(-strata)
#   kable(t, caption = name)
# }))
```

@@

- `tgroup=1` indicates the years 2003 - 2008
- `tgroup=2` indicates the years 2009 - 2013
- `tgroup=3` indicates the years 2014 - 2018
